The inefficacy of superficial similarities for improving instructor-student relationships

### 2021-04-20 outputs/paper/paper.Rmd
add april 19 draft

---
title: "The inefficacy of superficial similarities for improving instructor-student relationships"
author: "Amy Farrow"
date: "April 19th, 2021"
header-includes:
  \usepackage{dcolumn}
  \usepackage{rotating}
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "Replicating ‘Taking It to the Next Level’"
abstract: "This paper replicates the 2019 article 'Taking It to the Next Level' [@citeoriginal], which evaluates an intervention to improve college instructor-student relationship. The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity or instructor-student relationship, grades, or re-enrollment. While the scalability and affordability of the intervention are desirable, there are no results in any of the targeted measures: those that affect and reflect college retention. These results are consistent with the original paper.\\par \\textbf{Keywords:} instructor-student relationship, college, replication study"
thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
bibliography: references.bib
---

@@

@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests an intervention to improve college retention and performance. In this field experiment, they tested the effect of instructor-student similarity on instructor-student relationships (ISRs) and measures of student success. Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive result. The experiment consisted of a randomized controlled trial where some undergraduate students were informed of similarities they shared with their instructor, while others were not. Student and instructor perceptions of similarity and the instructor-student relationship were measured through surveys, and student performance measures were collected from school records.

@@

These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grades, or persistence. These conclusions stand when matching instructor-student traits are considered, and when vulnerable student groups are evaluated independent of the wider student population. The lack of significant effects reflects the complicated task of increasing perceived similarity, to improve instructor relationships, and to then improve student outcomes, especially in a college context that may make instructor-student similarity and connection challenging.

@@

```{r modeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(studentfirstsimilaritymodel,
          studentsecondsimilaritymodel,
          studentsecondrelationshipmodel,
          teachersecondrelationshipmodel,
          grademodel,
          examgrademodel,
          teachersecondsimilarity,
          enrollmentfallmodel,
          title = "Replication model results",
          align = TRUE,
          ci = TRUE,
          float.env = "sidewaystable",
          no.space = TRUE,
          keep.stat = c("rsq", "adj.rsq", "aic", "lr", "wald", "ll", "logrank"),
          #omit.stat=c("ser","f", "ll"),
          column.sep.width = "-10pt",
          label = "tab:modeltable",
          type = "latex",
          header = FALSE,
          initial.zero = FALSE,
          omit = c("teacherid"),
          omit.labels = c("Instructor fixed effect"),
          dep.var.labels = c("Student sim. 1",
                             "Student sim. 2",
                             "Student ISR 2",
                             "Instructor ISR 2",
                             "Course grade",
                             "Final grade",
                             "Instructor sim. 2",
                             "Persistence"),
          covariate.labels = c("Treatment", "Student ISR 1", "Student female", "CGPA"))
```

>>>>

---
title: "The inefficacy of superficial similarities for improving instructor-student relationships"
author: "Amy Farrow"
date: "April 19th, 2021"
header-includes:
  \usepackage{dcolumn}
  \usepackage{rotating}
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "Replicating ‘Taking It to the Next Level’"
abstract: "This paper replicates the 2019 article 'Taking It to the Next Level' [@citeoriginal], which evaluates an intervention to improve college instructor-student relationship. The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity or instructor-student relationship, grades, or re-enrollment. The goal of this intervention was to improve student performance and persistence, and while the scalability and affordability are desirable, there are no results in any of the targeted measures. These results are consistent with the original paper.\\par \\textbf{Keywords:} instructor-student relationship, college, replication study"
thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
bibliography: references.bib
---

@@

@citeoriginal's paper, "Taking It to the Next Level: A Field Experiment to Improve Instructor-Student Relationships in College", tests an intervention to improve college retention and performance. In this field experiment, they evaluated the relationships between perceived similarity, instructor-student relationships (ISRs), and measures of student success. Based on extensive K-12 research about the importance of instructor-student relationship for student success, @citeoriginal aimed to establish how instructor-student relationships could be improved at the college level, and to test if this improvement had a positive result. The experiment consisted of a randomized controlled trial where some undergraduate students were informed of similarities they shared with their instructor, while others were not. Student and instructor perceptions of similarity and the instructor-student relationship were measured through surveys, and student performance measures were collected from school records.

@@

These models show that treatment slightly improves student perception of instructor-student similarity, but does not significantly affect student or instructor ISR perception, grades, or persistence. These conclusions stand when matching instructor-student traits are considered and when vulnerable student groups are evaluated separately from the wider student population. The lack of significant effects reflects the complicated task of increasing perceived similarity, hopefully to improve instructor relationships, and to then improve student outcomes, especially in a college context that may make instructor-student similarity and connection challenging.

@@

```{r modeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(studentfirstsimilaritymodel,
          studentsecondsimilaritymodel,
          studentsecondrelationshipmodel,
          teachersecondrelationshipmodel,
          grademodel,
          examgrademodel,
          teachersecondsimilarity,
          enrollmentfallmodel,
          title = "Replication model results",
          align = TRUE,
          ci = TRUE,
          float.env = "sidewaystable",
          no.space = TRUE,
          omit.stat=c("ser","f"),
          column.sep.width = "-10pt",
          label = "tab:modeltable",
          type = "latex",
          header = FALSE,
          initial.zero = FALSE,
          omit = c("teacherid"),
          omit.labels = c("Instructor fixed effect"),
          dep.var.labels = c("Student sim. 1",
                             "Student sim. 2",
                             "Student ISR 2",
                             "Instructor ISR 2",
                             "Course grade",
                             "Final grade",
                             "Instructor sim. 2",
                             "Persistence"),
          covariate.labels = c("Treatment", "Student ISR 1", "Student female", "CGPA"))
```
